---
title: "Tex15_VNO_BulkRNAseq_Codes"
output: html_document
---

# setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = '', fig.width = 10, fig.height = 10)
```

```{r LoadPackages, message=FALSE}
library(biomaRt)
library(tidyverse)
library(plyranges)
library("rtracklayer")
library(ChIPQC)
library(Rsubread)
library(DESeq2)
library(ggrastr)
library(ggbreak)
library("RColorBrewer")
library(edgeR)
library(circlize) # circos plots and colors for complexheatmap
library(randomcoloR)
library(ComplexHeatmap)
library(patchwork)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(biomaRt)
library(tidyverse)
library(plyranges)
library("rtracklayer")
library(ChIPQC)
library(Rsubread)
library(DESeq2)
library(ggrastr)
library(ggbreak)
library("RColorBrewer")
library(edgeR)
library(circlize)
library(randomcoloR)
library(ComplexHeatmap)
library(patchwork)
library(naniar)
library(neonUtilities)
library(fgsea)
library(rstatix)
library(ggpubr)
library(vegan)
```
# Universal set up
```{r functions}
gr_to_bed <- function(gr){
gr %>% as_tibble() %>% dplyr::select(seqnames,start,end,name,score,strand)  
}
format_GRanges_for_featureCounts <- function(GRanges_intervals){
GRanges_intervals  %>% as_tibble() %>% dplyr::select("GeneID" = name,Chr=seqnames,Start=start,End=end,Strand=strand)
}
Get_GRanges_overlap <- function(Reference,Query,FieldName){
output <- bind_rows(data.frame(Reference[Reference %over% Query]) %>% mutate( !!FieldName := TRUE),data.frame(Reference[!(Reference %over% Query)]) %>% mutate( !!FieldName := FALSE)) %>% arrange(seqnames,start) %>% GRanges()
return(output)
}
save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
    stopifnot(!missing(x))
    stopifnot(!missing(filename))
    pdf(filename, width=width, height=height)
    grid::grid.newpage()
    grid::grid.draw(x$gtable)
    dev.off()
}
```

```{r GenomeAnnotation}
included_contigs <- c(1:19,"X","Y")
contig_levels <- str_c("chr",c(1:19,"X","Y"))
blklist <- import.bed("/seq/mm10/mm10-blacklist.v3.bed")


```

```{r GeneAnnotation}
gene_table <- read_delim(file="/media/storage2/kevin/paper_Tex15/annotation/ENSEMBL_plus_Barnes_with_Pcdh_constant_gene_reference.txt",delim="\t",col_names = TRUE)

# gene sets
ORs <- read_delim(file="/media/storage2/kevin/paper_Tex15/annotation/OR_annotation.txt",delim="\t",col_names = TRUE) %>% distinct() %>% dplyr::rename(cluster_name = name) %>% mutate(seqnames = factor(seqnames,levels=contig_levels)) %>% droplevels()
TypeI_ORs <- ORs %>% filter(class == "ClassI")
TypeII_ORs <-  ORs %>% filter(class == "ClassII") 

# gene sets
Olfr <- gene_table %>% semi_join(ORs) %>% mutate(receptor_class = "Olfr")
Olfr_I <- gene_table %>% semi_join(TypeI_ORs) %>% mutate(receptor_class = "Olfr_ClassI")
Olfr_II <- gene_table %>% semi_join(TypeII_ORs) %>% mutate(receptor_class = "Olfr_ClassII")
Taar <- gene_table %>% filter(grepl("Taar",gene_name)) %>% mutate(receptor_class = "Taar")
Vmn1r <- gene_table %>% filter(grepl("Vmn1",gene_name))  %>% mutate(receptor_class = "Vmn1r")
Vmn2r <- gene_table %>% filter(grepl("Vmn2",gene_name))  %>% mutate(receptor_class = "Vmn2r")
Fpr <- gene_table %>% filter(grepl("Fpr",gene_name)) %>% mutate(receptor_class = "Fpr")
ReceptorGenes <- bind_rows(list(Olfr_I,Olfr_II,Taar,Vmn1r,Vmn2r,Fpr))
NonReceptorGenes <-  gene_table %>% anti_join(ReceptorGenes)
ReceptorGenes_withNon <- NonReceptorGenes %>% mutate(receptor_class = "non_GPCR") %>% bind_rows(ReceptorGenes)

cPcdh <- gene_table %>% filter(grepl("Pcdha",gene_name) | grepl("Pcdhb",gene_name) | grepl("Pcdhg",gene_name) )
```

# Load Data

## VNO_Tex15
```{r}
# Load Data

## VNO Tex15 Data
VNO_Tex15_libraries.all <- read_delim(file="/media/storage3/nader/Tex15_VNO_Nov25_Libraries/VNO_Libraries_Sheet8.csv",delim=",",col_names = TRUE) %>% 
  mutate(across(c(Library,SampleName,Group,fastq_folder),factor)) %>% 
  mutate(Exp_check = c(Group=="VNO.Tex15KO"))%>%
  mutate(Exp_Group= if_else(Exp_check == TRUE,true = "Experimental",false = "Control")) %>%
  select(-Exp_check)

  
```

### Library QC
```{r figwidth = 10, message=FALSE,warning=FALSE,fig.width=8,fig.height=8}
### Library QC
p.VNO_Tex15_raw_reads <- ggplot(VNO_Tex15_libraries.all)+
	geom_bar(aes(x=SampleName,y=(raw_pairs/1000000),fill=Group),stat = "identity")+
	geom_hline(yintercept=20,linetype=2,color="black")+
	scale_y_continuous(expand = c(0, 0))+
	theme_bw()+
	theme(aspect.ratio=2,legend.position = "none",text = element_text(size=18),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
	labs(y="Raw read pairs (million)",x="",title="Sequencing depth")+
	coord_cartesian(ylim=c(0,140))
ggsave(p.VNO_Tex15_raw_reads,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/p.VNO_Tex15_raw_reads.pdf")
p.VNO_Tex15_raw_reads

p.VNO_Tex15_percent_unique <- ggplot(VNO_Tex15_libraries.all)+
	geom_bar(aes(x=SampleName,y=(percent_unique),fill=Group),stat = "identity")+
	scale_y_continuous(expand = c(0, 0))+
	theme_bw()+
	theme(aspect.ratio=2,legend.position = "none",text = element_text(size=18),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
	labs(y="Percent unique",x="",title="percent uniquely aligned")+
	coord_cartesian(ylim=c(0,100))
ggsave(p.VNO_Tex15_percent_unique,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/p.VNO_Tex15_percent_unique.pdf")
p.VNO_Tex15_percent_unique
```


### filter out low qualitry or overly shallow libraries
```{r message=FALSE, warning=FALSE}
### filter out low qualitry or overly shallow libraries
VNO_Tex15_libraries <- VNO_Tex15_libraries.all %>% mutate(bam_file_name=bamFile_all,bampath = bam_all_path) %>% filter(!grepl("fail",QC)) %>% filter(!grepl("poor",QC)) # select single end data
VNO_Tex15_metaData <- VNO_Tex15_libraries %>% dplyr::select(row_names=bam_file_name,Group,Exp_Group,Library,SampleName,bam_file=bam_file_name,fastq_folder) %>% column_to_rownames(var = "row_names") 
VNO_Tex15_bamsToCount <- VNO_Tex15_libraries %>%  pull(bampath)
```

```{r countReads}
# count reads
VNO_Tex15_counts  <- featureCounts(files=file.path(VNO_Tex15_bamsToCount),annot.ext="/media/working/seq/mm10/annotation/Monahan_Lab_gene_annotation_mm10_2025-02/Monahan_Lab_gene_annotation_mm10_2025-02.gtf" ,isPairedEnd=TRUE,isGTFAnnotationFile=TRUE,GTF.featureType="exon",GTF.attrType="gene_id",nthreads=12,strandSpecific=0,useMetaFeatures = TRUE)
saveRDS(VNO_Tex15_counts,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_counts.rData")
```


```{r runDESeq2}

VNO_Tex15_metaData<- VNO_Tex15_metaData %>% mutate(Exp_Group = ifelse(Group== "VNO.Tex15Het_Admera","Control","Experimental"))

VNO_Tex15_dds <- DESeqDataSetFromMatrix(countData = VNO_Tex15_counts$counts, colData = VNO_Tex15_metaData, design = ~  Exp_Group)
mcols(VNO_Tex15_dds) <- data.frame(basepairs = VNO_Tex15_counts$annotation$Length)
VNO_Tex15_dds <- VNO_Tex15_dds[ rowMeans(counts(VNO_Tex15_dds)) > 1,  ] # filter to remove low expressed genes
VNO_Tex15_dds$Group <- relevel(VNO_Tex15_dds$Group, ref = "VNO.Tex15Het_Admera")
VNO_Tex15_dds <- DESeq(VNO_Tex15_dds) # perform differential expression analysis
saveRDS(VNO_Tex15_dds,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_dds.rData")
```

```{r}
VNO_Tex15_counts <- readRDS(file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_counts.rData")
VNO_Tex15_dds <- readRDS(file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_dds.rData")
```




### Sample comparison and generating Figure 2B
```{r}
VNO_Tex15_vsd <- vst(VNO_Tex15_dds, blind=FALSE)
VNO_Tex15_rld <- rlog(VNO_Tex15_dds, blind=FALSE)
head(assay(VNO_Tex15_vsd), 3)

VNO_Tex15_sampleDists <- dist(t(assay(VNO_Tex15_vsd)))
VNO_Tex15_sampleDistMatrix <- as.matrix(VNO_Tex15_sampleDists)
rownames(VNO_Tex15_sampleDistMatrix) <- paste(VNO_Tex15_vsd$Group, VNO_Tex15_vsd$type, sep="-")
colnames(VNO_Tex15_sampleDistMatrix) <- NULL
VNO_Tex15_colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(VNO_Tex15_sampleDistMatrix,
         clustering_distance_rows=VNO_Tex15_sampleDists,
         clustering_distance_cols=VNO_Tex15_sampleDists,
         col=VNO_Tex15_colors)


#Generating figure 2B
VNO_Tex15_pcaData <- plotPCA(VNO_Tex15_vsd, intgroup=c("Group"), returnData=TRUE)
VNO_Tex15_percentVar <- round(100 * attr(VNO_Tex15_pcaData, "percentVar"))
PCA_plot<-ggplot(VNO_Tex15_pcaData, aes(PC1, PC2, color=Group)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",VNO_Tex15_percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",VNO_Tex15_percentVar[2],"% variance")) + 
  labs(labels= "Experimental Group", values= c("Control","Experimental"))+
  coord_fixed()+theme_bw()

ggsave(PCA_plot,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/PCA_plot.pdf",height=7,width=7)
```

### Generate counts from DEseq2
```{r get_gene_included_gene_names}
VNO_Tex15_fpkm <- fpkm(VNO_Tex15_dds, robust = FALSE) %>% as.data.frame() %>% rownames_to_column(var="gene_id") %>% pivot_longer(cols=c(-gene_id),values_to="fpkm",names_to="bam_file") %>% left_join(VNO_Tex15_metaData)
VNO_Tex15_included_genes <- VNO_Tex15_fpkm %>% pull(gene_id) 
```

```{r calculateTPMs}
# calculate TPMs by ENSEMBL id
VNO_Tex15_tpm <- VNO_Tex15_fpkm %>% group_by(Library) %>% mutate(total_fpkm = sum(fpkm)) %>% mutate(tpm = (fpkm / total_fpkm) *1000000) %>% dplyr::select(gene_id,Library,Group,tpm) %>% ungroup()


VNO_Tex15_tpm.table <- VNO_Tex15_tpm %>% dplyr::select(gene_id,Library,tpm) %>% pivot_wider(names_from="Library",values_from="tpm") %>% left_join(gene_table)
VNO_Tex15_tpm.mean <- VNO_Tex15_tpm %>% group_by(Group,gene_id) %>% summarize(mean_tpm = mean(tpm)) %>% arrange(gene_id) %>% ungroup()
VNO_Tex15_tpm.mean.table <- VNO_Tex15_tpm.mean %>% pivot_wider(values_from=mean_tpm,names_from=Group) %>% left_join(gene_table)

worked_ReceptorGenes<-ReceptorGenes%>% select(gene_id, receptor_class)


worked_VNO_Tex15_tpm<- left_join(VNO_Tex15_tpm.table, worked_ReceptorGenes, suffix = c(".x",".y"))

 write_delim(worked_VNO_Tex15_tpm, file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_tpm_full_11-25.csv",delim=",")

```


### Results
```{r}
### Results
experiment_name <- c("rna_VNO_Tex15KO-vs-Control")
Groups <- tibble(Group = c("Control","Experimental"))
selected_contrast <- c("Exp_Group", "Experimental","Control")
significance_threshold <- 0.05 

VNO_Tex15_res <- results(VNO_Tex15_dds, alpha=significance_threshold, contrast= selected_contrast)
VNO_Tex15_results <- VNO_Tex15_res %>% as.data.frame() %>% rownames_to_column(var="gene_id") %>% mutate(significant = ifelse((padj< significance_threshold),"TRUE","FALSE")) %>% replace_na(list(significant = "FALSE"))  %>% mutate(direction = ifelse(log2FoldChange < 0, "down","up")) %>% left_join(gene_table)

write_delim(VNO_Tex15_results,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_results_11-25.csv",delim=",")

ReceptorGenes %>% inner_join(VNO_Tex15_results) %>% write_delim(file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_results_Receptors_11-25.csv",delim=",")
```

#DESeq2 results table
```{r}
#DESeq2 results table
VNO_Tex15KO_res <- results(VNO_Tex15_dds, alpha=significance_threshold, contrast= c("Exp_Group", "Experimental","Control"))
VNO_Tex15KO_results <- VNO_Tex15KO_res %>% as.data.frame() %>%
  rownames_to_column(var="gene_id") %>%
  mutate(significant = ifelse((padj< significance_threshold),"TRUE","FALSE")) %>%
  replace_na(list(significant = "FALSE"))  %>%
  mutate(direction = ifelse(log2FoldChange < 0, "down","up"),probability = padj) %>%
  mutate(set=ifelse(significant==TRUE,direction,"NS"),probability=padj) %>%
  left_join(gene_table)
write_delim(VNO_Tex15KO_results ,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15KO_results.csv",delim=",")

```


#DESeq2 results table, lfcThreshold = 1.5
```{r}
#DESeq2 results table, lfcThreshold = 1.5
VNO_Tex15KO_res_lfc_Threshold <- results(VNO_Tex15_dds, alpha=significance_threshold, contrast= c("Exp_Group", "Experimental","Control"),lfcThreshold = log2(1.5))
VNO_Tex15KO_results_lfc_Threshold <- VNO_Tex15KO_res_lfc_Threshold %>% as.data.frame() %>%
  rownames_to_column(var="gene_id") %>%
  mutate(significant = ifelse((padj< significance_threshold),"TRUE","FALSE")) %>%
  replace_na(list(significant = "FALSE"))  %>%
  mutate(direction = ifelse(log2FoldChange < 0, "down","up"),probability = padj) %>%
  mutate(set=ifelse(significant==TRUE,direction,"NS"),probability=padj) %>%
  left_join(gene_table)
write_delim(VNO_Tex15KO_results_lfc_Threshold ,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15KO_results_lfc_Threshold.csv",delim=",")

```


```{r}
resultsNames(VNO_Tex15_dds)
selected_contrast = "Exp_Group_Experimental_vs_Control"

VNO_Tex15_resShrink <- lfcShrink(VNO_Tex15_dds, coef=selected_contrast, type="ashr") 
VNO_Tex15_resultsShrink <- VNO_Tex15_resShrink %>% as.data.frame() %>% rownames_to_column(var="gene_id") %>% mutate(significant = ifelse((padj< significance_threshold),"TRUE","FALSE")) %>% replace_na(list(significant = "FALSE"))  %>% mutate(direction = ifelse(log2FoldChange < 0, "down","up")) %>% left_join(gene_table)
write_delim(VNO_Tex15_resultsShrink,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_resultsShrink_11-25.csv",delim=",")


```

#Generate tables for change up or change down
```{r}
#Generate tables for change up or change down
VNO_Tex15_DEG_Up <- VNO_Tex15_results  %>% filter(padj < 0.05) %>% filter(log2FoldChange > 0)
VNO_Tex15_DEG_Down <- VNO_Tex15_results  %>% filter(padj < 0.05) %>% filter(log2FoldChange < 0)
```


#Generating the main TPM table from DEseq2 and other information
```{r}
#Generating the main TPM table from DEseq2 and other information

##Reads the TPM Table
tpmtable<- read.delim(file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_tpm_full_11-25.csv", sep = ",")
##Create Control Mean
tpmtable$Hetmean <- rowMeans(tpmtable[c("rna.VNO.Tex15HET.6416.41825","rna.VNO.Tex15HET.6349.42525","rna.VNO.Tex15HET.7135.101725","rna.VNO.Tex15HET.7137.101725")], na.rm = TRUE)
##Create Experimental Means
tpmtable$KOmean<- rowMeans(tpmtable[c("rna.VNO.Tex15KO.7136.101725","rna.VNO.Tex15KO.7131.101725","rna.VNO.Tex15KO.7132.101725","rna.VNO.Tex15KO.7134.102125")], na.rm = TRUE)
## Creating TPM delta between control and experimental
tpmtable$TPM_Delta <- tpmtable$Hetmean- tpmtable$KOmean

##adding statistics to the table
stat_Table<-read.delim(file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/VNO_Tex15_resultsShrink_11-25.csv", sep = ",")

Part1_Complete_Table<- inner_join(tpmtable,stat_Table)

##pulling V2R annotations
V2R_complete_Family<-read_delim(file="/media/storage3/nader/Tex15_VNO_KO_tpm/V2Rs groups by Family - V1 Sheet.csv",delim=",",col_names = TRUE)
Vmn2r_annotation <-read_delim(file="/media/storage2/kevin/paper_Tex15/annotation/Vmn2r_annotation.csv",delim=",",col_names = TRUE)
Vmn2r_annotation_modded <- Vmn2r_annotation %>% select(-V2R_Family)
Complete_Vmn2r_annotation<- left_join(Vmn2r_annotation_modded,V2R_complete_Family, by= c("gene_name"))

##pulling V1R annotations
Vmn1r_cluster_annotation<-read_delim(file="/media/storage2/kevin/paper_Tex15/annotation/Vmn1r_annotation.csv", delim=",",col_names = TRUE)
###Standardising V1R cluster names
SingchromV1R<- Vmn1r_cluster_annotation %>% filter(nchar(seqnames)==4)
SingchromV1R.2<-SingchromV1R %>% mutate("cluster_name"= substr(SingchromV1R$cluster_name, start = 6, stop = 20))
duochromV1R<- Vmn1r_cluster_annotation %>% filter(nchar(seqnames)==5)
duochromV1R.2<-duochromV1R%>% mutate("cluster_name"= substr(duochromV1R$cluster_name, start = 7, stop = 20))
Vmn1r_cluster_annotation_final<- full_join(duochromV1R.2,SingchromV1R.2)

Vmn1r_subfamilies<-read_delim(file="/media/storage3/nader/Tex15_VNO_KO_tpm/V1R_Subfamilies.csv", delim=",",col_names = TRUE)
Vmn1r_annotation <- left_join(Vmn1r_cluster_annotation_final, Vmn1r_subfamilies, by= c("gene_name", "gene_biotype"))

##merging V1R and V2R cluster annotations
Vr_cluster_annotation <- full_join( Vmn1r_cluster_annotation_final,Complete_Vmn2r_annotation)


##Generating final table for processing
One_Table_trta<- left_join(Part1_Complete_Table,Vr_cluster_annotation, by= c("gene_id","gene_name","gene_biotype"))

write_delim(One_Table_trta,file =  "/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/One_Table_trta_Final.csv",append = FALSE, delim = ",")

```


<font size="5"> *Generating filter tables from non pivoted TPM tables*

```{r}
#Generating Filter Tables

One_Table_trta<- read.delim(file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/One_Table_trta_Final.csv", sep = ",")

##Fpr Table
FPR_table<-One_Table_trta %>% filter(receptor_class== "Fpr")

##C family V2Rs
V2R_CFam_list <-One_Table_trta %>% filter(cluster_name== "Vmn2r_cluster_2")

##A, B, and D family V2Rs
V2R_NCFam_list<-One_Table_trta %>% filter(receptor_class== "Vmn2r") %>% anti_join(V2R_CFam_list) 

##V1Rs
V1R_list<-One_Table_trta %>% filter(receptor_class== "Vmn1r")


```


#Generating the geom jitter of the VR clusters for Figure S2A and B
```{r}

#Generating the geom jitter of the VR clusters for Figure S2A and B


##pulling the cluster names of the V2Rs and replacing NA values with "Unknown"
V2Rclusternames<-  V2R_NCFam_list %>% select(cluster_name) %>% distinct(cluster_name) %>% mutate("cluster_name"= ifelse(is.na(cluster_name),"Unknown",cluster_name))%>%arrange(cluster_name)%>% pull(1)  

##Factoring the names of the V2R clusters
V2R_NCFam_list$cluster_name<-factor(V2R_NCFam_list$cluster_name, levels = V2Rclusternames , labels=c("Vmn2r_cluster_8", "Vmn2r_cluster_9","Vmn2r_cluster_10", "Vmn2r_cluster_11", "Vmn2r_cluster_12", "Vmn2r_cluster_13", "Vmn2r_cluster_14", "Vmn2r_cluster_15", "Vmn2r_cluster_19", "Vmn2r_cluster_20", "Vmn2r_cluster_22", "Vmn2r_cluster_24","Vmn2r_cluster_25", "Vmn2r_cluster_26", "Vmn2r_cluster_30", "Vmn2r_cluster_31","Vmn2r_cluster_39", "Vmn2r_cluster_40", "Vmn2r_cluster_42", "Vmn2r_cluster_44", "Vmn2r_cluster_45","Vmn2r_cluster_46","Unknown")) 

##S2B
V2R_NCFam_list_boxplot<- V2R_NCFam_list %>% ggplot()+
  theme_bw()+
  geom_jitter( aes(x= cluster_name, y = log2FoldChange, colour= significant),alpha=.5, width = .1)+ 
  ggtitle("Log2fold Change in A, B, and D Family V2R Clusters") + 
  xlab("Cluster")+
  scale_x_discrete(labels=c("Cluster 8","Cluster 9","Cluster 10","Cluster 11","Cluster 12","Cluster 13","Cluster 14","Cluster 15", "Cluster 19","Cluster 20","Cluster 22","Cluster 24","Cluster 25","Cluster 26","Cluster 30","Cluster 31","Cluster 39","Cluster 40","Cluster 42"," Cluster 44","Cluster 45","Cluster 46","Unknown"))+
  labs(fill= "Gene")+
  ylab("Log2fold Change")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(aspect.ratio = 1)
V2R_NCFam_list_boxplot

##S2A
V1R_listboxplot<- V1R_list  %>% arrange(cluster_name) %>% ggplot()+
  theme_bw()+
  geom_jitter( aes(x= cluster_name, y = log2FoldChange, colour= significant),alpha=.5, width = .1)+ 
  ggtitle("Log2fold Change in V1R Clusters") + 
  xlab("Cluster")+ 
  scale_x_discrete(labels= c("Cluster 1","Cluster 2","Cluster 3","Cluster 4","Cluster 5","Cluster 6","Cluster 7"))+
  labs(fill= "Gene")+
  ylab("Log2fold Change")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(aspect.ratio = 2)
V1R_listboxplot

ggsave(V2R_NCFam_list_boxplot,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V2R_NCFam_list_boxplot.pdf",height=7,width=7)
ggsave(V1R_listboxplot,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V1R_listboxplot.pdf",height=7,width=7)

```

##Generating Figure 2C
```{r}

Trusig_table <-  One_Table_trta %>% replace_na(list(receptor_class = "non_receptor")) %>% filter(receptor_class== "Vmn1r"|receptor_class== "Vmn2r"|receptor_class== "Fpr"|receptor_class== "non_receptor") %>%  group_by(receptor_class,direction, significant) %>% summarise(count=n()) %>% ungroup()%>% group_by(direction, significant) %>% mutate("Percentage"=(count/sum(count))*100)

Trusig_column_up<- Trusig_table%>%filter(significant== TRUE)%>% filter(direction=="up") %>% ggplot(aes(x=direction, y= Percentage, fill = receptor_class))+geom_col()+ theme_bw()+theme(aspect.ratio = 2)+
  scale_fill_manual(values=c("Fpr"="royalblue","non_receptor"="springgreen2","Vmn1r"="goldenrod1","Vmn2r"="darkorchid2"))+ coord_polar("y",start=0)+theme_void()
Trusig_column_up

Trusig_column_down<- Trusig_table%>%filter(significant== TRUE)%>% filter(direction=="down") %>% ggplot(aes(x=direction, y= Percentage, fill = receptor_class))+geom_col()+ theme_bw()+theme(aspect.ratio = 2)+
  scale_fill_manual(values=c("Fpr"="royalblue","non_receptor"="springgreen2","Vmn1r"="goldenrod1","Vmn2r"="darkorchid2"))+ coord_polar("y",start=0)+theme_void()
Trusig_column_down

Falsesig_table <-  One_Table_trta %>% replace_na(list(receptor_class = "non_receptor")) %>% filter(receptor_class== "Vmn1r"|receptor_class== "Vmn2r"|receptor_class== "Fpr"|receptor_class== "non_receptor") %>% filter(significant==FALSE) %>%  group_by(receptor_class, significant) %>% summarise(count=n()) %>% ungroup()%>% mutate("Percentage"=(count/sum(count))*100)

Faslesig_column<- Falsesig_table %>% ggplot(aes(x=significant, y= Percentage, fill = receptor_class))+geom_col()+  coord_polar("y",start=0) +theme_bw()+theme(aspect.ratio = 2)+
  scale_fill_manual(values=c("Fpr"="royalblue","non_receptor"="springgreen2","Vmn1r"="goldenrod1","Vmn2r"="darkorchid2"))+theme_void()
Faslesig_column



ggsave(Trusig_column_up,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/piechart_sigup.pdf",height=7,width=7)
ggsave(Trusig_column_down,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/piechart_sigdown.pdf",height=7,width=7)
ggsave(Faslesig_column,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/piechart_ns.pdf",height=7,width=7)


```
#Generating Tables for Figure S2E S2F S2A
```{r}



##generating base table
Box_worked_One_Table_trta<-One_Table_trta %>% dplyr::select(-KOmean,-Hetmean)  %>% pivot_longer(cols = c(rna.VNO.Tex15HET.6416.41825,rna.VNO.Tex15HET.7135.101725,rna.VNO.Tex15KO.7136.101725, rna.VNO.Tex15HET.7137.101725,rna.VNO.Tex15KO.7131.101725,rna.VNO.Tex15KO.7134.102125,rna.VNO.Tex15KO.7132.101725,rna.VNO.Tex15HET.6349.42525))%>% mutate("Exp_Group"=ifelse(grepl("HET",name),"Control","Experimental"))

###Generating H2-MV table
worked_H2MV_list_box<- Box_worked_One_Table_trta %>% filter(is.na(receptor_class)) %>% filter(grepl("H2-M",gene_name)) 

###Generating Vmn2r81 table
worked_vmn2r81_list <-Box_worked_One_Table_trta %>% filter(receptor_class== "Vmn2r")  %>% filter(grepl("Vmn2r81",gene_name))

###Generating Vmn2r82 table
worked_vmn2r82_list <-Box_worked_One_Table_trta %>% filter(receptor_class== "Vmn2r")  %>% filter(grepl("Vmn2r82",gene_name))

###Generating Joined Vmn2r81 and Vmn2r82 table
worked_vmn2r8182_list <-full_join(worked_vmn2r82_list,worked_vmn2r81_list)

##Generating Figure 2F
vmn2r8182_boxplot<- worked_vmn2r8182_list  %>% ggplot()+
  theme_bw()+
  geom_boxplot(aes(x= gene_name, y = value, fill= Exp_Group),width = 0.4, outlier.shape = NA)+
  geom_point( aes(x= gene_name, y = value, fill= Exp_Group),width=0.05, position= position_dodge(width = .4))+ 
  ggtitle("TPM Comparison between Control and Knockout") + 
  xlab("Gene")+
  labs(fill= "Experimental Group")+
  ylab("Transcripts Per Million (TPM)")+ theme(aspect.ratio = 1)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  coord_cartesian(ylim=c(0,50)) 
vmn2r8182_boxplot

##Generating Figure 2G
H2mv_boxplot<- worked_H2MV_list_box%>%  filter(significant==TRUE)%>% ggplot()+
  theme_bw()+
  geom_boxplot(aes(x= gene_name, y = value, fill= Exp_Group),width = 1, outlier.shape = NA)+
  geom_jitter( aes(x= gene_name, y = value, fill= Exp_Group), position= position_dodge(width = 1))+ 
  ggtitle("TPM Comparison between Control and Knockout") + 
  xlab("Gene")+
  labs(fill= "Experimental Group")+
  ylab("Transcripts Per Million (TPM)")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(aspect.ratio = 1)+
  coord_cartesian(ylim=c(0,70)) 
H2mv_boxplot


#Total tpm table made
worked_VR_list_box<- Box_worked_One_Table_trta %>%filter(receptor_class== "Vmn1r"|receptor_class== "Vmn2r")

sum_worked_VR_list_box_notsig <-worked_VR_list_box %>% group_by(name,receptor_class)%>% mutate(sumtpm= sum(value)) %>% distinct(name,Exp_Group,sumtpm)%>% arrange(Exp_Group)
t.test(sumtpm~Exp_Group, data=sum_worked_VR_list_box_notsig)
sum_worked_VR_list_box_notsig

#Figure S2A
worked_VR_sum_comparison_box <- sum_worked_VR_list_box_notsig %>%
  ggplot()+
  theme_bw()+
  geom_boxplot(aes(x= receptor_class, y = sumtpm, fill = Exp_Group),width = 0.4, outlier.shape = NA)+
  geom_point( aes(x= receptor_class, y = sumtpm, fill = Exp_Group),width=0.05, position= position_dodge(width = .4))+coord_cartesian(ylim=c(0,2000))+ theme(aspect.ratio = 1)
worked_VR_sum_comparison_box


ggsave(worked_VR_sum_comparison_box,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/sum_tpm_boxplot.pdf",height=7,width=7)
ggsave(H2mv_boxplot,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/H2mv_boxplot.pdf",height=7,width=7)
ggsave(vmn2r8182_boxplot,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/vmn2r8182_boxplot.pdf",height=7,width=7)

```

#Generating Line plots for Figure 2D 2E 2F S2B
```{r}


##figure 2D
V1R_list_Line <- V1R_list %>% mutate(set = ifelse(significant == TRUE,direction,"NS")) %>% ggplot() + aes(log10(Hetmean+1),log10(KOmean+1), fill = set, colour = set) +geom_point(size=2, shape= 23)+geom_abline(yintercept=0,color="orange")+
  ggtitle("V1R Family Expression")+ 
  coord_fixed(ratio = 1)+ 
  coord_cartesian(xlim= c(0,3), ylim=c(0,3))+
  coord_fixed(ratio = 1, xlim = c(0,3),ylim = c(0,3))+
  labs(fill= "Transcription Change")+ 
  theme_bw() + 
  xlab("Log 10 Control Mean")+
  ylab("Log 10 Knockout Mean")+
  #scale_colour_manual(values = c("indianred4","springgreen4"))+
  scale_fill_manual(values=c("NS"="black","up"="blue","down"="red"))+
  scale_colour_manual(values=c("NS"="black","up"="blue","down"="red"))
V1R_list_Line

##figure 2E
V2R_NCfam_Line <- V2R_NCFam_list%>% mutate(set = ifelse(significant == TRUE,direction,"NS")) %>% ggplot() + aes(log10(Hetmean+1),log10(KOmean+1), fill = set, colour = set) +geom_point(size=2, shape= 23)+geom_abline(yintercept=0,color="orange")+
  ggtitle("V2R ABD Family Expression")+ 
  coord_fixed(ratio = 1)+ 
  coord_cartesian(xlim= c(0,3), ylim=c(0,3))+
  coord_fixed(ratio = 1, xlim = c(0,3),ylim = c(0,3))+
  labs(fill= "Transcription Change")+ 
  theme_bw() + 
  xlab("Log 10 Control Mean")+
  ylab("Log 10 Knockout Mean")+
  #scale_colour_manual(values = c("indianred4","springgreen4"))+
  scale_fill_manual(values=c("NS"="black","up"="blue","down"="red"))+
  scale_colour_manual(values=c("NS"="black","up"="blue","down"="red"))
V2R_NCfam_Line

##Figure 2F
V2R_Cfam_Line <- V2R_CFam_list%>% mutate(set = ifelse(significant == TRUE,direction,"NS")) %>% ggplot() + aes(log10(Hetmean+1),log10(KOmean+1), fill = set, colour = set) +geom_point(size=2, shape= 23)+geom_abline(yintercept=0,color="orange")+ ggtitle("V2R C Family Expression")+ coord_fixed(ratio = 1)+ coord_cartesian(xlim= c(0,3), ylim=c(0,3))+coord_fixed(ratio = 1, xlim = c(0,3),ylim = c(0,3))+labs(fill= "Transcription Change")+ theme_bw() + 
  xlab("Log 10 Control Mean")+
  ylab("Log 10 Knockout Mean")+
  scale_fill_manual(values=c("NS"="black","up"="blue","down"="red"))+
  scale_colour_manual(values=c("NS"="black","up"="blue","down"="red"))
V2R_Cfam_Line 

##Figure S2B
Fpr_Line <- FPR_table %>% mutate(set = ifelse(significant == TRUE,direction,"NS")) %>% ggplot() + aes(log10(Hetmean+1),log10(KOmean+1), fill = set, colour = set) +geom_point(size=2, shape= 23)+geom_abline(yintercept=0,color="orange")+ ggtitle("Fpr Expression")+ coord_fixed(ratio = 1)+ coord_cartesian(xlim= c(0,1), ylim=c(0,1))+coord_fixed(ratio = 1, xlim = c(0,1),ylim = c(0,1))+labs(fill= "Transcription Change")+ theme_bw() + 
  xlab("Log 10 Control Mean")+
  ylab("Log 10 Knockout Mean")+
  scale_fill_manual(values=c("NS"="black","up"="blue","down"="red"))+
  scale_colour_manual(values=c("NS"="black","up"="blue","down"="red"))
Fpr_Line

ggsave(V1R_list_Line ,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V1R_list_Line.pdf", width= 7.2917, height=4.4861)
ggsave(V2R_Cfam_Line,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V2R_Cfam_line.pdf", width= 7.2917, height=4.4861)
ggsave(V2R_NCfam_Line ,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V2R_NCfam_Line.pdf", width= 7.2917, height=4.4861)
ggsave(Fpr_Line,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/Fpr_Line.pdf", width= 7.2917, height=4.4861)

```

#Ratio ranking Figure 2G 2H
```{r}

worked_One_Table_trta<-One_Table_trta %>% dplyr::select(-rna.VNO.Tex15HET.6416.41825
,-rna.VNO.Tex15HET.6349.42525
,-rna.VNO.Tex15HET.7135.101725
,-rna.VNO.Tex15KO.7136.101725
, -rna.VNO.Tex15HET.7137.101725
, -rna.VNO.Tex15KO.7131.101725, -rna.VNO.Tex15KO.7134.102125,
-rna.VNO.Tex15KO.7132.101725	)  %>% pivot_longer(cols = c(-gene_id,-gene_name,-gene_biotype,-receptor_class,-TPM_Delta,-baseMean,-log2FoldChange,-lfcSE,-pvalue,-padj,-significant,-direction,-seqnames,-start,-end,-width,-strand,-score,-cluster_name,-V2R_Family,-V2R_Type))

worked_V2R_CFam_list <- worked_One_Table_trta %>% filter(cluster_name== "Vmn2r_cluster_2")

worked_V2R_NCFam_list <-worked_One_Table_trta %>% filter(receptor_class== "Vmn2r") %>% anti_join(worked_V2R_CFam_list)

worked_V1R_list<-worked_One_Table_trta %>% filter(receptor_class== "Vmn1r")

V1R_receptor_rank <- worked_V1R_list %>% mutate(name = factor(name,levels=c("Hetmean","KOmean"))) %>% dplyr::select(name,value) %>% 
    group_by(name) %>% arrange(name,-value) %>%
    mutate(rank = row_number()) %>% mutate("Percentage"=(value/sum(value)))%>%
  arrange(-Percentage) %>% 
  mutate(cumulative_fraction=cumsum(Percentage)) %>%
  ungroup()%>% arrange(name, rank)

V1R_receptor_rank_graph<-V1R_receptor_rank %>% 
  ggplot()+geom_line(aes(x=rank,y=cumulative_fraction,group=name,color=name))+
  scale_y_continuous(labels = scales::percent) + 
  theme_classic()+
	labs(y="Cumulative Percentage",x="V1R expression Rank",title="V1R Table")+ theme(aspect.ratio = 1)
V1R_receptor_rank_graph

V2R_receptor_rank <- worked_V2R_NCFam_list %>% mutate(name = factor(name,levels=c("Hetmean","KOmean"))) %>% dplyr::select(name,value) %>% 
    group_by(name) %>% arrange(name,-value) %>%
    mutate(rank = row_number()) %>% mutate("Percentage"=(value/sum(value)))%>%
  arrange(-Percentage) %>% 
  mutate(cumulative_fraction=cumsum(Percentage)) %>%
  ungroup()%>% arrange(name, rank)
   
V2R_receptor_rank_graph<-V2R_receptor_rank %>% 
  ggplot()+geom_line(aes(x=rank,y=cumulative_fraction,group=name,color=name))+
  scale_y_continuous(labels = scales::percent) + 
  theme_classic()+
	labs(y="Cumulative Percentage",x="V2R expression Rank",title="V2R Table")+ theme(aspect.ratio = 1)
V2R_receptor_rank_graph

ggsave(V2R_receptor_rank_graph,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V2R_receptor_rank_graph.pdf")
ggsave(V1R_receptor_rank_graph,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/V1R_receptor_rank_graph.pdf")

```

#Shannon's Index Diversity Significance Calculation Figure 2I
```{r}

VR_list<-One_Table_trta %>% filter(receptor_class== "Vmn1r"|receptor_class== "Vmn2r")

divtablenomean <- VR_list%>% select(gene_name, rna.VNO.Tex15HET.6416.41825
,rna.VNO.Tex15HET.6349.42525
,rna.VNO.Tex15HET.7135.101725
, rna.VNO.Tex15HET.7137.101725,rna.VNO.Tex15KO.7136.101725, rna.VNO.Tex15KO.7131.101725, rna.VNO.Tex15KO.7134.102125,
rna.VNO.Tex15KO.7132.101725)

rownames(divtablenomean) <- divtablenomean$gene_name
divtablenomean$gene_name <- NULL
  
divtablenomeant<-t(divtablenomean)

divnomean<-diversity(divtablenomeant)

divnomean_tibble <- as_tibble(divnomean, rownames = "Sample")%>%
  mutate("Experimental_Group"=ifelse(grepl("HET",Sample),"Control","Experimental"))%>% 
  group_by(Experimental_Group)%>% 
  mutate("Shannon_Index_mean"=mean(value))%>% 
  rename(value="Shannon_Index")
divnomean_tibble

result<-t.test(Shannon_Index~Experimental_Group, divnomean_tibble)
p.value<-result$p.value

p.divnomean_tibble<- divnomean_tibble  %>%
      mutate(p_value = p.value)

div_plot<- p.divnomean_tibble %>% ggplot(aes(x=Experimental_Group, y=Shannon_Index))+theme_bw()+ geom_dotplot(binaxis = "y",stackdir = "center", binwidth = .01)+ theme(aspect.ratio = 1)+stat_compare_means(aes(x= Experimental_Group, y = Shannon_Index), method = "t.test", label = "p.format", label.y= 5.5, label.x = 1.4)
div_plot

ggsave(div_plot,file="/media/storage3/nader/Tex15_VNO_Preprint_Code/Bulk_RNAseq_Figures/div_plot.pdf")

write_delim(p.divnomean_tibble,file =  "/media/storage3/nader/Tex15_VNO_Preprint_Code/Generated_files/Shannon Index Table.csv",append = FALSE, delim = ",")


```
